### 2011-11-17 longRange.hetero/writeDown/main.tex
add main.tex

\documentclass[aps,pre,preprint]{revtex4-1}
% \documentclass[aps,pre,twocolumn]{revtex4-1}
% \documentclass[aps,jcp,groupedaddress,twocolumn,unsortedaddress]{revtex4}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[dvips]{graphicx}
\usepackage{color}
\usepackage{tabularx}

\makeatletter
\makeatother

\newcommand{\recheck}[1]{{\color{red} #1}}
\renewcommand{\v}[1]{\textbf{\textit{#1}}}
\renewcommand{\d}[1]{\textsf{#1}}

\begin{document}

\title{The Error Estimate of Force Calculation in the Inhomogeneous Molecular Systems}
\author{Han Wang}
\affiliation{LMAM and School of Mathematical
  Sciences, Peking University}
\author{Pingwen Zhang}
% \email{pzhang@pku.edu.cn}
\affiliation{LMAM and School of Mathematical
  Sciences, Peking University}

\begin{abstract}
\end{abstract}

\maketitle

\section{Theory}
\subsection{Ewald summation}
The Ewald method divides the electrostatic interaction in to three
parts: the direct part, the reciprocal part and the correction
part. that is
\begin{equation}
E = E_{\textrm{dir}} + E_{\textrm{rec}} + E_{\textrm{corr}},
\end{equation}
where 
\begin {align}
E_{\textrm{dir}} & = \frac12 \sum^{\ast}_{\v n}
\sum_{i,j = 1}^{N} \frac{q_iq_j \textrm{erfc}(\beta \vert\v{r}_{ij} + \v{n}\vert)}
{\vert\v{r}_{ij} + \v{n}\vert} \\
E_{\textrm{rec}} & = \frac1{2\pi V} \sum_{\v m \neq 0}
\frac{\exp(-\pi^2\v m^2 / \beta^2)}{\v m^2} S(\v m) S(-\v m) \\
 E_{\textrm{corr}}& = -\frac\beta{\sqrt \pi} \sum_{i=1}^N q_i^2
\end {align}
The distance between the two particles is denoted by $\v r_{ij} = \v
r_i - \v r_j$, where $\v r_i$ and $\v r_j$ are positions of particle
$i$ and $j$, respectively.  The lattice in the real space is denoted
by $\v n = n_1 \v a_1 + n_2 \v a_2 + n_3 \v a_3$, where $\v a_\alpha,
\ \alpha=1,2,3$ are box vectors. the structure factor $S(\v m)$ is
defined by
\begin{equation}\label{sm1}
S(\v m) = \sum_{j=1}^N q_j \exp (2 \pi i \v m \cdot \v r_j),
\end{equation}
where $\v m = m_1 \v a_1^\ast + m_2 \v a_2^\ast + m_3 \v a_3^\ast$ is
the reciprocal space lattice vectors. $\v a_\alpha^\ast,\ \alpha = 1,
2, 3$ are conjugate reciprocal vectors of $\v a_\alpha$, defined by
$\v a_\alpha \cdot \v a_\beta^\ast = \delta_{\alpha\beta}$. and $V =
\v a_1 \cdot \v a_2 \times \v a_3$ is the volume of the box.

Both the direct and the reciprocal parts are calculated by cut-off in
real and reciprocal space, respectively. For the reciprocal part the
cut-offed energy reads
\begin{align}
  E_{\textrm{rec}} & =
  \frac1{2\pi V}
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \frac{\exp(-\pi^2\v m^2 / \beta^2)}{\v m^2} S(\v m) S(-\v m)   
\end{align}
where $K_\alpha$ is the number of Fourier modes used on direction
$\alpha$. For convenience, we denote
\begin{align}
  f(\v m) = \frac{\exp(-\pi^2 \v m^2 / \beta^2)}{2\pi V \v m^2}
\end{align}
The truncated reciprocal force acting on particle $i$ is
\begin{align}
  \v F(\v r_i) = - 
  q_i 
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m) \,
  S(-\v m)\,
  e^{2\pi i\v m\cdot\v r_i}
\end{align}
where
\begin{align}
  \v g(\v m) = -4 \pi \v m i f(\v m)
\end{align}

\subsection{Error estimate for the reciprocal Ewald force}

In this paper, we consider the force error, bacause the applicaions
are based on the molecular dynamics simulation. We study the force
error at arbitrary position $\v r$ by adding a testing particle with
charge $q$ to $\v r$. The difference between the exact force and
calculated force (with error introduced by cut-offs) exerting on the
testing particle are denoted by $\v F(\v r)$ and $\tilde{\v F}(\v r)$,
respectively. Note that the testing particle does NOT exert any force
on the system. The difference between the exact and calculated is
called the \emph{error force} \cite{short}, and is denoted by $\Delta
\v F(\v r) = \v F(\v r) - \tilde{\v F}(\v r)$.  The error of the force
computation is defined by the \emph{root mean square (RMS)}
\emph{error}, which is the square root of the second moment of the
error force: $\mathcal E(\v r) = \sqrt{\langle\vert\Delta\v F(\v
  r)\vert^2\rangle}$.  The notation $\langle\cdot\rangle$ means
ensemble average.

Ref.~\cite{wang2010optimizing} pointed out that it
is reasonable and convenient to estimate the error for the direct and
reciprocal force part seperately. 
The direct part is a short-range
interaction in real space, the error estimate of which
is well established in \cite{short}.
The mean of the error force is
\begin{align}\label{eqn:tmp10}
  \langle\Delta\v F_{\textrm{dir}}(\v r)\rangle
  =q\, [\v K^c_{\textrm{dir}}\ast\rho_q](\v r).
\end{align}
Herein, the ``$\ast$'' notation means convolution.  The direct
convolutional kernel of the error force $\v K^c_{\textrm{dir}}(\v r)$
is defined by
\begin{align}
  \v K^c_{\textrm{dir}}(\v r) =
  \left\{
    \begin{aligned}
      &\,0, &\quad & \vert\v r\vert \leq r_c;\\
      &\frac{\textrm{erfc}(\beta\vert\v r\vert)}{\vert\v r\vert},& & \vert\v r\vert > r_c,
    \end{aligned}
  \right.
\end{align}
where $r_c$ is the cut-off radius in the real space.  The first order
charge distribution function $\rho_q$ is defined by:
\begin{align}
  \rho_q(\v r) = 
  \bigg\langle
  \sum_{j = 1}^N
  \,q_j\,\delta(\v r - \v r_j)
  \bigg\rangle.
\end{align}
With the help of fast Fourier transform (FFT), the convolution in
\eqref{eqn:tmp10} can be calculated with the computational cost of
$\mathcal O(N\log N)$.

And the RMS error is estimated by
\begin{align}\nonumber
  \langle\vert\Delta\v F_{\textrm{dir}}(\v r)\vert^2\rangle
  = &\,
  q^2\,[(\v K^c_{\textrm{dir}})^2\ast\rho_{q^2}] (\v r) + 
  q^2\,[\v K^c_{\textrm{dir}}\ast\rho_{q}]^2 (\v r) + \\
  &
  q^2\,\int_{\mathbb R^3\times\mathbb R^3}
  \v K^c_{\textrm{dir}}(\v r-\v r')\cdot
  \v K^c_{\textrm{dir}}(\v r-\v r'')\,
  C_{q^2}(\v r', \v r'')\,
  \d d\v r'\d d\v r''.
\end{align}
The three terms on the right hand side are homogeneity error,
inhomogeneity error and correlation error. The second order charge
distribution $\rho_{q^2}$ is
\begin{align}
  \rho_{q^2}(\v r) = 
  \bigg\langle
  \sum_{j = 1}^N
  \,q^2_j\,\delta(\v r - \v r_j)
  \bigg\rangle.
\end{align}
The charge density correlation function is defined by
\begin{align}
  C_{q^2}(\v r', \v r'')
  =
  \bigg\langle
  \sum_{i\neq j}q_iq_j\delta(\v r'-\v r_i)\delta(\v r''-\v r_j)
  \bigg\rangle
  - \rho_q(\v r')\rho_q(\v r'').
\end{align}
In most situations, neglecting the correlation error will give
satisfactory estimate of the error~\cite{short}. Therefore, 
\begin{align}
  \langle\vert\Delta\v F_{\textrm{dir}}(\v r)\vert^2\rangle
  = &\,
  q^2\,[(\v K^c_{\textrm{dir}})^2\ast\rho_{q^2}] (\v r) + 
  q^2\,[\v K^c_{\textrm{dir}}\ast\rho_{q}]^2 (\v r).
\end{align}

The reciprocal error force, namely the difference between the
the exact and truncated reciprocal force, is
\begin{align}
  \Delta \v F_{\textrm{rec}}(\v r)
  = & - 
  q
  \sum_{\vert m_\alpha\vert \geq K_\alpha/2}
  \v g(\v m) \,
  S(-\v m)\,
  e^{2\pi i\v m\cdot\v r} 
\end{align}
It is proved that the mean reciprocal error force is
\begin{align}
  \big\langle
  \Delta \v F_{\textrm{rec}}(\v r)
  \big\rangle
  &= 
  -q\,[\v K_c\ast\rho_q\,] (\v r_i),
\end{align}
where the reciprocal convolutional kernel of the error force $\v
K^c_{\textrm{rec}}$ is defined by:
\begin{align}
  \v K^c_{\textrm{rec}}(\v r) =
  \sum_{
      \vert m_\alpha\vert \geq K_\alpha/2}
  \v g(\v m) \,
  e^{2\pi i\v m\cdot\v r}.
\end{align}
Similarly, with the assumption of uncorrelation, it is shown that
\begin{align}
  \langle\vert\Delta\v F_{\textrm{rec}}(\v r)\vert^2\rangle
  = &\,
  q^2\,[(\v K^c_{\textrm{rec}})^2\ast\rho_{q^2}] (\v r) + 
  q^2\,[\v K^c_{\textrm{rec}}\ast\rho_{q}]^2 (\v r).
\end{align}

From \eqref{eqn:tmp13}, the ensemble average of the squared force error is
\begin{align}\nonumber
  \big\langle
  \vert\Delta\v F(\v r_i)\vert^2
  \big\rangle
  = \,&
  q^2_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{1,\alpha}\vert \geq K_\alpha/2\\
      \v m_1\neq 0
    \end{subarray}}
  \sum_{
    \begin{subarray}{c}
      \vert m_{2,\alpha}\vert \geq K_\alpha/2\\
      \v m_2\neq 0
    \end{subarray}}
  \v g(\v m_1)\cdot\v g(\v m_2) \,
  e^{2\pi i(\v m_1 + \v m_2)\cdot\v r_i} \\ \nonumber
  &
  \bigg\langle
  \sum_{j\neq i}q_j\,e^{-2\pi i\v m_1\cdot\v r_j}  
  \sum_{k\neq i}q_k\,e^{-2\pi i\v m_2\cdot\v r_k}
  \bigg\rangle\\ \nonumber
  = \,&
  q^2_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{1,\alpha}\vert \geq K_\alpha/2\\
      \v m_1\neq 0
    \end{subarray}}
  \sum_{
    \begin{subarray}{c}
      \vert m_{2,\alpha}\vert \geq K_\alpha/2\\
      \v m_2\neq 0
    \end{subarray}}
  \v g(\v m_1)\cdot\v g(\v m_2) \,
  e^{2\pi i(\v m_1 + \v m_2)\cdot\v r_i}\\\nonumber
  &
  \bigg\langle
  \sum_{j\neq i}\,q_j^2\,e^{-2\pi i(\v m_1 + \v m_2)\cdot\v r_j} +
  \sum_{j\neq k}\,q_jq_k\, e^{-2\pi i\v m_1\cdot\v r_j} e^{-2\pi i\v m_2\cdot\v r_k}
  \bigg\rangle
\end{align}
Assuming particle $i$ is independent with $j$ when $i\neq j$, the second term is
approximately $\langle\Delta\v F(\v r_i)\rangle^2$, therefore,
\begin{align}\nonumber
  \big\langle
  \vert\Delta\v F(\v r_i)\vert^2
  \big\rangle
  = \,&
  q^2_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{1,\alpha}\vert \geq K_\alpha/2\\
      \v m_1\neq 0
    \end{subarray}}
  \sum_{
    \begin{subarray}{c}
      \vert m_{2,\alpha}\vert \geq K_\alpha/2\\
      \v m_2\neq 0
    \end{subarray}}
  \v g(\v m_1)\cdot\v g(\v m_2) \\\nonumber
  \,&
  \int
  \bigg\langle
  \sum_{j\neq i}\,q_j^2
  \delta (\v r' - \v r_j)
  \bigg\rangle
  \,e^{2\pi i(\v m_1 + \v m_2)\cdot(\v r_i - \v r')}
  \,\d d\v r' \\ \nonumber
  &+\,
  \langle\Delta\v F(\v r_i)\rangle^2 \\ \nonumber
  =\,&
  q^2_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{1,\alpha}\vert \geq K_\alpha/2\\
      \v m_1\neq 0
    \end{subarray}}
  \sum_{
    \begin{subarray}{c}
      \vert m_{2,\alpha}\vert \geq K_\alpha/2\\
      \v m_2\neq 0
    \end{subarray}}
  \v g(\v m_1)\cdot\v g(\v m_2) 
  \int
  \,e^{2\pi i(\v m_1 + \v m_2)\cdot(\v r_i - \v r')}
  \,\rho_{q^2}(\v r')
  \,\d d\v r' \\  \nonumber
  &+\,
  \langle\Delta\v F(\v r_i)\rangle^2 \\ 
  =\;&
  q^2_i
  \int \v G^2_c(\v r - \v r')\,\rho_{q^2}(\v r')\,\d d\v r'  
  + 
  \langle\Delta\v F(\v r_i)\,\rangle^2 \\
  =\;&
  q_i^2\,[\v G_c^2\ast\rho_{q^2}] (\v r_i) + \langle\Delta\v F(\v r_i)\,\rangle^2 
\end{align}

% \newpage

\bibliography{ref}{}
\bibliographystyle{unsrt}

\end{document}