### 2012-04-20 shortRange.inhomo/alltext/slides/ucsb.2012.04/slide.tex
add ucsb

\documentclass{beamer}

% \usepackage{CJKnumb}\usepackage{beamerthemesplit}

\mode<article>
{
  \usepackage{beamerbasearticle}
  \usepackage{fullpage}
  \usepackage{hyperref}
}

%\usepackage{beamerthemesplit} 
%\usepackage{beamerthemeshadow}  
%\usepackage[width=2cm,dark,tab]{beamerthemesidebar}

% Setup appearance:

%\usetheme{Darmstadt}
\usefonttheme[onlylarge]{structurebold}
\setbeamerfont*{frametitle}{size=\normalsize,series=\bfseries}
\setbeamertemplate{navigation symbols}{}

\renewcommand\arraystretch{1.5}

% Standard packages

\usepackage[english]{babel}
%\usepackage[latin1]{inputenc}

\usepackage{epsf}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{tabularx}

% \usepackage[usenames,dvipsnames]{color}
\definecolor{shadow}{gray}{0.8}
\newcommand{\redc}[1]{{\color{red} #1}}
\newcommand{\bluec}[1]{{\color{blue} #1}}
\newcommand{\shadowc}[1]{{\color{shadow} #1}}
\definecolor{myyellow}{HTML}{FFB700}
\newcommand{\yellowc}[1]{{\color{myyellow} #1}}
\newcommand{\greenc}[1]{{\color{green} #1}}
\renewcommand{\v}[1]{\textbf{\textit{#1}}}
\renewcommand{\d}[1]{\textrm{#1}}

\usepackage{amsfonts}
\newcommand{\tickYes}{\checkmark}
\usepackage{pifont}
\newcommand{\tickNo}{\hspace{1pt}\ding{55}}

% \usetheme{Boadilla}
% \usetheme{Copenhagen}
% \usetheme{Madrid}
\usetheme{Singapore}

\begin{document}
%\title{Comparative atomistic and coarse-grained study of water: Simulation details vs. Simulation feasibility}
%\title[Optimizing SPME]{Optimizing Working Parameters of the Smooth Particle Mesh Ewald Algorithm}
\title[]{
  Estimating the force computation error in inhomogeneous and correlated
  molecular systems
}
%
\author{Han Wang}
\institute[FUB] {
  Institute for Mathematics, Freie Universit\"at Berlin, Germany
\vskip 0.4cm
Joint with: Christof Sch\"utte (FUB), Pingwen Zhang (PKU)}
\date[12 Apr 2012]{12 Apr 2012}
\frame{\titlepage}

% \begin{frame}{Why?}
%   \begin{itemize}
%     \vfill
%   \item <1-> Why force computation?
%     \begin{itemize}\itemsep 3pt
%     \item <2->Molecular dynamics simulation.
%     \item <3->Computationally intensive: \redc{$ \sim 90 \%$}
%     \item <4->One of the main sources of the \redc{``error''}.
%     \item <5->May introduce \redc{unphysical artifacts}.
%     \end{itemize}
%     \vfill
%   \item <6-> Why error estimate?
%     \begin{itemize}\itemsep 3pt
%     \item <7->Understanding \& correcting
%       the unphysical artifacts, quantitatively.
%     \item <8->Correction to force, energy, pressure, etc.
%     \item <9->Boost the efficiency of simulation automatically.
%     \end{itemize}
%     \vfill
%   \end{itemize}
% \end{frame}

\begin{frame}{Why?}
  Why force computation?
  \begin{itemize}\itemsep 3pt
  \item <2->Molecular dynamics simulation.
  \item <3->Computationally intensive: \redc{$ \sim 90 \%$}
  \item <4->One of the main sources of the \redc{``error''}.
  \item <5->May introduce \redc{unphysical artifacts}.
  \end{itemize}
  \vfill
\end{frame}

\begin{frame}{Unphysical artifacts due to the force error}
  \begin{itemize}\itemsep 10pt
  \item Example system:
    \begin{figure}
      \centering
      \includegraphics[width=0.9\textwidth]{figs/t0.85-n16000-rc07.5uni/confout-02.eps}
    \end{figure}
  \item Equilibrium densities:
    \begin{figure}
    \centering
    \includegraphics[width=0.49\textwidth]{figs/converge.new/t0p85-gas-1.eps} 
    \includegraphics[width=0.49\textwidth]{figs/converge.new/t0p85-liquid-1.eps} 
  \end{figure}  
  \end{itemize}
\end{frame}

\begin{frame}{Why?}
  Why error estimate?
  \begin{itemize}\itemsep 3pt
  \item <2->Understanding \& correcting
    the unphysical artifacts, quantitatively.
  \item <3->Correction to \redc{force, energy, pressure, etc.}
  \item <4->Boost the accuracy \& efficiency of simulation automatically.
  \end{itemize}  
  \vfill
\end{frame}

\begin{frame}{The working parameters of SPME} {A 6 dimensional parameter space!}
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/long-range/error-uniform.eps}
  \end{figure}  
\end{frame}

\begin{frame}{Overview of the error estimates}
  \begin{itemize}
  \vfill
  \item<1->   Conditions for the error estimate:
  \begin{enumerate}\itemsep 3pt
  \item \redc{Homogeneity}.
  \item Particles are \redc{uncorrelated}.
  \end{enumerate}
  \vfill
\item<2->   Current error estimates:
  \begin{table}
    \centering
    \begin{tabular*}{0.85\textwidth}{l@{\extracolsep{\fill}}ll}\hline\hline
      Conditions & cut-off & SPME \\\hline
      1+2 & \bluec{\tickYes\quad$\mathcal O(1)$}  & \bluec{\tickYes\quad$\mathcal O(1)$} \\
      2   & \redc{\tickYes\quad$\mathcal O(N\log N)$} & \redc{\tickYes\quad$\mathcal O(N\log N)$} \\
      none& \redc{\tickNo\quad$\mathcal O(N^2\log N)$} & \redc{\tickNo\quad$\mathcal O(N^2\log N)$} \\
          &  & \redc{N.N.A. $\mathcal O(N\log N)$} \\\hline\hline
    \end{tabular*}
  \end{table}
  \vfill
  \end{itemize}
\end{frame}

\begin{frame}{The error estimate}
  \begin{itemize}
  \item <1-> The error force
    \bluec{
      \begin{align*}
        \Delta \v F(\v r)
        &=
        \v F_{\textrm{calculate}}(\v r) - \v F_{\textrm{exact}}(\v r)
      \end{align*}}
  \item <2-> Define the error force kernel
    \bluec{
      \begin{align*}
        \Delta \v F(\v r)
        &=
        q \sum_{j=1}^N q_j \,\v K(\v r, \v r_j)
      \end{align*}}
    It is possible for most main stream force calculation algorithms: cut-off, SPME, P3M, etc.
  \item <3-> The kernel of cut-off method:
    \bluec{
      \begin{align*}
        \v K(\v r,\v r') =
        \left\{
          \begin{array}{ll}
            0, & \vert\v r - \v r'\vert\leq r_c; \\
            \v F(\vert\v r-\v r'\vert), & \vert\v r - \v r'\vert > r_c.
          \end{array}
        \right.
      \end{align*}}
    \vfill
  \end{itemize}
\end{frame}

\begin{frame}{The error estimate}
  \begin{itemize}\itemsep -10pt
  \item<1->   The mean error force is
    \bluec{
      \begin{align*}
        \langle\Delta\v F(\v r)\rangle
        =
        q\int_{\mathbb R^3}\v K(\v r , \v r')\,\rho_q(\v r')\,\d d\v r'
      \end{align*}
    }
  \vskip -10cm
\item<2->   The error:
  \bluec{
    \begin{align*} \nonumber
      \langle\vert\Delta\v F(\v r)\vert^2\rangle
      = 
      \redc{\mathcal E^2_{\textrm{homo}}(\v r)} +
      % \langle\Delta\v F(\v r)\,\rangle^2 +
      \redc{\mathcal E^2_{\textrm{inhomo}}(\v r)} +
      \redc{\mathcal E_{\textrm{correlation}}(\v r)}.
    \end{align*}
  }
  with\bluec{
  \begin{align*}
    \redc{\mathcal E^2_{\textrm{homo}}(\v r)}
    = &\,
    q^2\int_{\mathbb R^3}\vert\v K(\v r, \v r')\vert^2\rho_{q^2}(\v r')\,\d d\v r'  \\
    \redc{\mathcal E^2_{\textrm{inhomo}}(\v r)}
    = &\,
    q^2\bigg[\int_{\mathbb R^3}\v K(\v r, \v r')\rho_q(\v r')\,\d d\v r'\,\bigg]^2
    \redc{ = \langle\Delta\v F(\v r)\,\rangle^2}
    \\
    \redc{\mathcal E_{\textrm{correlation}}(\v r)}
    =&\,
    q^2\int_{\mathbb R^3\times\mathbb R^3}\v K(\v r, \v r')\cdot\v K(\v r, \v r'')\,C_{q^2}(\v r', \v r'')\,\d d\v r'\d d\v r''
  \end{align*}}
  \end{itemize}
\end{frame}

\begin{frame}{Liquid-vapor equilibrium of Lennard-Jones particles}
  \begin{figure}
    \centering
    \includegraphics[scale=1]{figs/t0.85-n16000-rc07.5uni/confout-02.eps}\\
    \includegraphics[]{figs/t0.85-n16000-rc07.5uni/error-uniform.eps}
  \end{figure}
\end{frame}

\begin{frame}{cut-off in bulk}
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/t0.85-n16000-rc07.5uni/cut-bulk-step00.eps}
  \end{figure}  
\end{frame}

\begin{frame}{cut-off in bulk}
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/t0.85-n16000-rc07.5uni/cut-bulk-step01.eps}
  \end{figure}  
\end{frame}

\begin{frame}{cut-off in bulk}
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/t0.85-n16000-rc07.5uni/cut-bulk-step02.eps}
  \end{figure}  
\end{frame}

\begin{frame}{cut-off in bulk}
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/t0.85-n16000-rc07.5uni/cut-bulk-step03.eps}
  \end{figure}  
\end{frame}

\begin{frame}{cut-off at interface}
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/t0.85-n16000-rc07.5uni/cut-bulk-step00.eps}
  \end{figure}  
\end{frame}

\begin{frame}{cut-off at interface}
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/t0.85-n16000-rc07.5uni/cut-interface-step03.eps}
  \end{figure}  
\end{frame}

\begin{frame}{cut-off at interface}
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/t0.85-n16000-rc07.5uni/cut-interface-step04.eps}
  \end{figure}  
\end{frame}

\begin{frame}{cut-off at interface}
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/t0.85-n16000-rc07.5uni/cut-interface-step05.eps}
  \end{figure}  
\end{frame}

\begin{frame}{Equilibrium gas density}
  \begin{figure}
    \centering
    \includegraphics[]{figs/converge.new/t0p85-gas.eps} 
  \end{figure}
  URC: uniform cut-off, ARC: adaptive cut-off, LFC: long-range force correction.
\end{frame}

\begin{frame}{Equilibrium liquid density}
  \begin{figure}
    \centering
    \includegraphics[]{figs/converge.new/t0p85-liquid.eps} 
  \end{figure}
  URC: uniform cut-off, ARC: adaptive cut-off, LFC: long-range force correction.
\end{frame}

\begin{frame}{Surface tension}
  \begin{figure}
    \centering
    \includegraphics[]{figs/converge.new/tension-t0p85.eps} 
  \end{figure}
  URC: uniform cut-off, ARC: adaptive cut-off, LFC: long-range force correction,
  LPC: long-range pressure correction.
\end{frame}

\begin{frame}{Electrostatic interaction and Ewald summation}
  \begin{itemize}
  \item<1->   Electrostatic interaction:
  \begin{equation*}\bluec{
    U_{\textrm{ele}} = \frac12 \sum^\ast_{n}\sum_{i,j}\frac{q_i q_j}{\vert \v r_{ij} + \v n\vert}}
  \end{equation*}
\item <2->
  Ewald summation (1921) splits the
  energy into three parts
  \bluec{
  \begin {align*}
    U_{\textrm{ele}} &=  U_{\textrm{dir}} + U_{\textrm{rec}}+ U_{\textrm{corr}}\\
    U_{\textrm{dir}} & = \frac12 \sum^{\ast}_{\v n}
    \sum_{i,j = 1}^{N} \frac{q_iq_j\, \redc{\textrm{erfc}(\beta \vert\v{r}_{ij} + \v{n}\vert)}}
    {\vert\v{r}_{ij} + \v{n}\vert} \\ \label{Erec-ewald}
    U_{\textrm{rec}} & = \frac1{2\pi V} \sum_{\v m \neq 0}
    \frac{\redc{\exp(-\pi^2\v m^2 / \beta^2)}}{\v m^2} S(\v m) S(-\v m) \\
    U_{\textrm{corr}}& = -\frac\beta{\sqrt \pi} \sum_{i=1}^N q_i^2
  \end {align*}}
  \end{itemize}
\end{frame}

\begin{frame}{Smooth Particle Mesh Ewald Method (SPME)}{Basic concepts}
  \begin{itemize}\itemsep -5pt
  \item<1-> The structure factor \bluec{$S(\v m)$} is
    defined by \bluec{
      \begin{equation*}\label{sm1}
        S(\v m) = \sum_{j=1}^N q_j \exp (2 \pi i \v m \cdot \v r_j),
      \end{equation*}}
  \item<2-> Optimal computational cost\footnote{
    \bluec{J. Perram, H. Petersen and S. De Leeuw, Mol. Phys. \textbf{65}, 875 (1988).}}: \redc{$\mathcal O(N^{1.5})$}.
    \vskip .5cm
  \item<3-> \redc{SPME}\footnote{
    \bluec{U. Essmann, \textit{et. al.}, J. Chem. Phys. \textbf{103}, 8577 (1995).}}: interpolate \bluec{$S(\v m)$} on grid, then FFT.
    The computational cost is 
  \begin{align*}
    \begin{tabular}[t]{c}
      \redc{$\mathcal O(N)$}\\
      {direct}
    \end{tabular}
    +
    \begin{tabular}[t]{c}
      \redc{$\mathcal O(N)$}\\
      {interpolation}
    \end{tabular}
    +
    \begin{tabular}[t]{c}
      \redc{$\mathcal O(N\log N)$}\\
      {FFTs}
    \end{tabular}
    =
    \begin{tabular}[t]{c}
      \redc{$\mathcal O(N\log N)$}\\
      {total cost}
    \end{tabular}
    \end{align*}
  \end{itemize}
\end{frame}

\begin{frame}{Smooth Particle Mesh Ewald Method (SPME)}
  {The working parameters is 6 dimensional}
  \vfill
  To use SPME method, one should provide the following parameters.
  \vfill
  \begin{itemize}
    \item \redc{$\beta$} \quad The controlling parameter.
    \item \redc{$r_c$} \quad The direct space cut-off radius.
    \item \redc{$\v K$} \quad The reciprocal space cut-off. The number of FFT grid points.
    \item \redc{$n$} \quad The order of cardinal B-spline interpolation.
  \end{itemize}
  \vfill An arbitrary combination may lead to \redc{totally wrong
    results}. \vfill
\end{frame}

\begin{frame}{The parameter tuning of SPME }
  The parameter tuning is a constrained optimization problem:\footnote{
    \bluec{H. Wang, F. Dommert and C. Holm, J. Chem. Phys. \textbf{133}, 034117 (2010).}}
  \bluec{
    \begin{align*} 
      \min\quad &  T (r_c, \v K, n, \beta),\\
      \textrm{\textbf{s.t.}}\quad & \mathcal E (r_c, \v K, n, \beta) = \mathcal E_{\textrm{C}}
    \end{align*}}
  \redc{Nearly optimal} working parameters can be found \redc{automatically}.
\end{frame}

\begin{frame}{The inhomogeneity problem of SPME}
  \begin{itemize}
  \item <1-> The mean error force is
    \bluec{
      \begin{align*}
        \langle\Delta\v F(\v r)\rangle
        =
        q\int_{\mathbb R^3}\v K(\v r , \v r')\,\rho_q(\v r')\,\d d\v r'
      \end{align*}
    }
    with
    \begin{align*}
      \bluec{  \rho_q(\v r) = 
        \bigg\langle
        \sum_{j = 1}^N
        q_j\delta(\v r - \v r_j)
        \bigg\rangle
      }
    \end{align*}    
  \item <2-> Varnishes if the system is \redc{locally neutral}.
  \end{itemize}
\end{frame}

\begin{frame}{Error estimate in an inhomogeneous and uncorrelated system}
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/long-range-inhomo/rand1-error.eps}
  \end{figure}  
\end{frame}

\begin{frame}{The trouble in water system}{Staggered mesh SPME}
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/long-range-inhomo/water-st-error-1.eps}
  \end{figure}  
\end{frame}

\begin{frame}{Nearest neighbor approximation of the correlation error}{Staggered mesh SPME}
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/long-range-inhomo/water-st-error-2.eps}
  \end{figure}  
\end{frame}

\begin{frame}
  \vfill
  \centerline{ \Huge
    Thanks!  }
  \vfill
\end{frame}

\end{document}