### 2012-07-09 longRange.hetero/alltext/publish/resubmit/lett.tex
back up

\documentclass[a4paper]{article}

\usepackage[dvips]{graphicx}

\begin{document}

\noindent
Reviewer  1 Comments:\\

\textit{ At a glance, the paper by Han Wang and colleagues appears to
  be another error estimator for P3M and related electrostatics
  methods.  But, after taking in the second paragraph it becomes clear
  that the authors are much more than mathematically inclined on the
  subject, and have indeed performed a useful service to the
  simulations community by incorporating the correlation of molecular
  charges into their error estimates.  It is probably not of
  tremendous practical benefit in an application like an Ewald
  parameter optimizer, as these programs can just repeat an
  electrostatic calculation on a representative snapshot of a
  simulation and estimate the error directly from the resulting
  forces.  However, it is of value to have a formal expression for the
  effect of charge correlations on these errors, and it may be
  instrumental in devising a pairwise approximate correction to
  electrostatic forces obtained from coarse grids.  The paper will
  require minor modifications so that the authors do not run afoul of
  various members of the community who hold strong opinions, but
  otherwise appears to be fit for publication.
}

?

\textit{
In the introduction, the authors should probably take the perspective that SPME is a special case of P3M.  In the opinions of the developers of SPME, the derivation differs from P3M although it leads to similar results, but even TA Darden acknowledges that the reciprocal space component of the non-smooth PME algorithm was inspired by P3M.  It is probably best to phrase SPME as such and not offend anyone.  The interlacing technique was first applied in the 1970s, as the authors have cited.  It was originally a means of reducing the memory requirements of these calculations—electrostatics on the two grids were solved serially.  Finally, the sentence “We prove that the error is decomposed…” should be changed to “We show that the error may be approximately decomposed…” to indicate that the estimator is, in the end, approximate but relevant to real systems.
}\\

not done\\

\textit{ Cosmetically, the authors should combine the two figures
  showing the charge distributions of Examples 1 and 2, to ensure they
  appear side-by-side in the published version.
}

We delete Fig.3 and plot the charge distribution of Example 2 together
with Example 1 in the new Fig.1. Changes to the caption and
descriptions are maded correspondingly.\\

\textit{
The authors are reasonable to focus on the error arising from the
reciprocal space approximations in the various P3M-related methods;
the error arising from the direct space part of the calculation is
simpler to assess.  However, because the authors give so much
attention to SPME and the Staggered Mesh method which, as published,
built on SPME, the authors should include some plot of the effect of
grid spacing, Gaussian smoothing function width (Ewald coefficient),
and interpolation order -- these are the determinants of the accuracy of
the reciprocal space calculation in these fast methods.  How does the
estimator do in terms of getting the various trends correct?  What is
the effect of correlated charge positions on the error and the
estimates under changing interpolation order?  This is at the heart of
the difference between ik- and analytic differentiation that the
authors have otherwise carefully investigated.  It would also be
helpful to see a plot of the error estimator’s performance as a
function of the “near neighbor” definition.  Define near neighbors
strictly by length, and plot the error estimator's accuracy as a
continuous function of the near neighbor cutoff.  Extra credit for
detailing the computation time as a function of the same parameter!
}\\

\textit{
One other point of discussion that might be relevant to this paper is
the significance of the results on force-shifted coulomb interactions
(see the simulation methods of Valerie Dagget and colleagues, or the
Wolfe Method as reviewed in J. Chem. Phys. 124:234104 and elsewhere).
These methods, based strictly on local “electrostatic” interactions
computed by modifying Coulomb’s law, are often trotted out as
alternatives to Ewald summation though their application is still
limited (which is, in the opinion of this reviewer, as it should be).
A principal assumption of these methods, which is often used
circuitously to defend their application, is that real systems are
locally neutral or that electrostatic correlations decay on the length
scale of a few water layers.  Wang and colleagues may be able to
analyze these methods, provide insights into the appearance of their
validity, and identify their limitations rigorously.  A formal
analysis of these methods is not required for this paper, but a
possible direction the authors might explore in the future.
}

We gratefully thank the reviewer for the suggestion of an interesting
extension of the present work.\\

\textit{
The authors have done an excellent job of manipulating the
mathematical foundations of these fast electrostatics methods.  With
minor modifications and extensions of the study, which this reviewer
does not intend to be onerous, this work should be published without
further hindrance.
}\\

\noindent
Reviewer  2 Comments:\\

\textit{ The article deals with the derivation of an accurate error
  estimate for the Ewald and SPME mesh Ewald method that can be used
  for inhomogeneous and correlated charged systems.
}

\textit{
Error estimates for long range algorithms are always useful for
periodic charged systems, since due to the long range nature, no exact
calculation is possible, but one rather has to truncate the
interactions at some point, and hence is forced to introduce errors.
Ideally, one would like to know for a specific method before doing an
expensive calculation how to choose method parameters such that I
reach a certain accuracy. Such error estimates have been developed in
the past, but normally only under the assumption that the charges are
randomly distributed. Here, the authors extend these error estimates
to the very relevant cases of inhomogeneous and/or correlated systems.
}

\textit{
  I find the article in general very readable and informative, and
  also suited for this journal. I have just a few comments that the
  authors might want to take into account in a minor revised version,
  before this article, in my opinion, is ready for publication:
}\\

Page 7, line 23:
The definition of a discrete set of nearest neighbors $\Omega_j$ might not readily avaible in all cases. For example, for simple liquids it may be more feasible to use the radial distribution function $g(r)$ up to a certain radius $r_c$, where $g(r) \approx 1$ for $r > c_c$. Then the sum has to be replaced by an integral over $r$.  Because of the close relation of $\rho^{(2)}$ and $g(r)$, the equations should be easy to adapt to such a notation.

Page 15, line 45:
I believe I saw that the for interlaced P$^3$M algorithm, which is very similar to the staggered SMPE algorithm,  the approximation to leading order of the error estimate is not justified, at least for higher precisions. This approximation could lead to a visible deviation from the true errors. In this context it might be useful to recalculate the numerical examples at higher precision. Usually, deviations start to be visible from about an RMS of $10^{-4}$.

Page 16, line 34:
Again, In the very similarly structured interlaced P$^3$M algorithm,  there are differences in accuracy between the $i\mathbf{k}$- and analytical differentiation at higher precisions.At this point it might not be justified  to assume that both can be treated in the same way. Similarly to the approximation of the $i\mathbf{k}$ error formula, it would be useful to do calculations at higher precision than those presented in the paper, and compare them

The manuscript contains a few misprints, and the language could be polished on a few occasions by a native speaker.

\vskip 1cm
\noindent
Reviewer 3 Comments:\\

\textit{
The authors have introduced a way to study electrostatic force errors
in correlated systems by assuming that the dominant correlations are
short-ranged which should usually work--they acknowledge this may not
work for systems near phase transitions. I like this paper and would
like to see it published after minor complaints are addressed.
}

\textit{
Firstly, the main contribution is to introduce a way to rapidly
evaluate the force error, including that due to short ranged (bonded)
pair interactions. Some readers may be confused by this, since bonded
pair electrostatic interactions are considered to be NOT
CALCULATED. However, in Ewald like methods the reciprocal contribution
cannot be easily excluded for atom pairs and is surprisingly large
even for bonded pairs. I think including this point somewhere in the
discussion will help many readers. Note that errors for bonded pairs
may be at least partly responsible for the disappointing performance
of multiple time step methods. (they are high frequency in time)
}

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{correlation.eps}
  \caption{The schematic plot of two interacting water molecules.
    When we calculate the force error of the left oxygen atom, the
    correlation with its bonded hydrogens are not considered, but
    the correlations within any neighboring molecule are considered,
    as indicated by the solid lines with two arrows.}
  \label{fig:1}
\end{figure}

We define the force error by inserting a testing charge into the
considered system, and comparing the calculated and exact force on
it. In the considered water system (Example 3), the subtle difference
between this setting to the realistic case is that the testing charge
has no correlation with the charges in the system, while any real
charge is bonded to two neighboring charges (via H-O bond and H-O-H
angle).  Therefore, the error estimate does not consider the
correlation between a charge and the its bonded neighbors, but it does
consider all bonding correlations within any neighboring molecule (see
Fig.~\ref{fig:1} of this letter). This is consistent with the usually
setting of bonded pair atoms, as mentioned by the reviewer, the
electrostatic interaction between them should not be calculated.

We therefore added the following sentence to make this point clearer:
\textbf{
  We define the force error by inserting a
testing charge into the water system, and comparing the
calculated and exact force on it. This definition implicitly assumes
the independency between the testing particle and the
system. Therefore, it sould be noted that the correlation error
estimate (20)
neglects the possible bonding
of the testing particle to any hydrogen/oxygen atom, and
calculates all bonded correlations
within its {neighboring} molecules. This is consistent
with the usual setting of the bonded atoms: the electrostatic interaction
between them is considered to be \emph{not} calculated.}

The inclusion of the correlation between the testing
charge and the system is also possible, but this is beyond the scope
of the present paper, and will be addressed in the following studies.\\

% Therefore, in all the estimates, the error due to the
% bonded pair interactions is automatically eliminated by the
% definition.

% The inclusion of the correlation between the testing
% charge and the system is also possible, but this is beyond the scope
% of the present paper, and will be addressed in the following studies.
% To make this point clearer, we add the following sentence:
% \textbf{Here we implicitly assumes that the testing particle is independent
% with the system, but it is not the case in real systems. For example,
% in the water system considered in Sec.III, each atom is bonded
% to two neighboring atoms (via the rigid H-O bond and H-O-H angle).
% Fortunately, this may not be a big problem, firstly because the bonded pair
% electrostatic interacions are considert to be \emph{not} calculated;
% secondly because the non-bonded pair correlation contributes only
% minority part to the total error (see also Sec.III).}

% As a matter of fact, in Example 3, the electrostatic interactions
% within one water molecule is \emph{not} calculated. This may not be
% computationally economic in practice, however, it is important to
% remove it in our theoretical study, becuase, as mensioned by the
% reviewer, the bonded pair electrostatic interaction is considered to
% be not calculated.

\textit{
I am curious about the fact that correlation reduces force errors
compared to usual estimates. I wonder if this is somehow due to
electrostatic forces on bonded pairs being anti-correlated due to
opposing charge signs e.g. on Oxygen and Hydrogens of $H_2O$.
}

As far as we concerned, it is not possible to test this point
directly, but we think it is a likely explanation to the observed
phenomenon.  Consider the extreme case: wihin one molecule, the
hydorgen and oxygen atoms locate at the same point (strongest
correlation), then the positive and negative charge cancels, so the
error vanishes. In case of TIP3P water model, the H-O distence is 0.1~nm and
the H-H distance is 0.16~nm, which are much smaller than the typical
center-of-mass distance between two neighboring molecules (0.31~nm).
We added this point to the paper.
\\

\textit{
Otherwise I have a number of minor grammatical and typo complaints:
}

The grammatical mistakes and typos have been corrected according to the reviewer.
% We want to explain two points:\\

% \textit{
% page 4 line 18 ``error estimate for the error force'' seems like 1st ``error'' is unnecessary
%               perhaps replace with ``following estimate for the error force''---also the term ``error force'' could be
%               replaced with ``force error'' here and elsewhere and I think the text would flow better
%             }

% We use ``error force'' because we want to stress it is a force, not a error, which may be a scalor, and be confused with the ``RMS force error''.\\

% \textit{
% page 24 line 11: ``the error is of good property'' Not clear what is meant. Do you mean total error is not
%                  contaminated by this?
%                }

%                Yes, the we removed ``the error is of good property'',
%                the logic of the sentence is

\end{document}